Convex Surrogates via Polyhedral Losses

### 2019-02-02 cvx-fin.tex
Moving stuff

\begin{corollary}\label{cor:d-embed-implies-d-construction}
  If a property is $d$-embeddable, then each of its level sets has a $d$-construction.
\end{corollary}
\begin{proof}
  Suppose the level set is represented as $\Gamma_r = \{p : Bp \geq 0\}$ with $B \in \reals^{k \times n}$ for the minimum $k$ (i.e. the number of facets of the polytope).
  Let $u \in \reals^d$ be its embedding point.
  The convex expected loss is minimized at $u$ if and only if it has the zero vector as a subgradient at $u$, i.e.
    \[ p \in \Gamma_r \qquad \iff \qquad 0 \in \partial \langle L(u), p \rangle . \]
  Subgradient sets satisfy the following two properties: $\partial(f_1 + f_2) = \partial f_1 + \partial f_2$ where the right side is the Minkowski sum of the subgradient sets (i.e. the set of sums of all pairs); and $\partial(\alpha f_1) = \alpha \partial f_1$.
  So
    \[ \partial \langle p , L(u) \rangle = \sum_{y \in \Y} p_y \partial L(u)_y \]
  where the right side is a Minkowski sum of subgradient sets.
  We now consider the halfspaces defining this set.
  Define the $y$th support function as $H_y(v) = \max_{w \in \partial L(u)_y} \langle w, v \rangle$ to obtain that $w \in \partial \langle p, L(u) \rangle$ if and only if, for all unit vectors $v$,
   \[ \langle v, w \rangle \leq \sum_{y \in \Y} p_y H_y(v) . \]
  In particular, the embedding point is optimal for $p$ if and only if this is satisfied for $w=0$.
  So, letting $B_i$ be the $i$th column of $B$, we have
    \[ \Gamma_r = \{p : \sum_y p_y B_y \geq 0\} = \{p : \sum_y p_y H_y(v) \geq 0 ~ (\forall v, \|v\|=1)\} . \]
  These sets are equal and full-dimensional as polytopes.
  By standard results (e.g.~\cite{grunbaum2013convex}), their H-representations are unique, so we must have $k$ vectors $v_1,\dots,v_k \in \reals^d$ such that $H_y(v_i) = B_{iy}$.
  Collected as rows of $V$, we obtain that each $(V,B_y)$ is a tight halfspace representation, as it contains $\partial L(u)_y$ and $H_y(v_i) = B_{iy}$, meaning there exists $w \in \partial L(u)_y$ such that $\langle w, v_i \rangle = B_{iy}$.
\end{proof}
This provides the following recipe for both constructing losses and proving lower bounds.
First, take any level set and represent it as $\{p : Bp \geq 0\}$.
Now if $B$ has $k$ rows, search for a collection of $k$ vectors in $\reals^d$.
For any candidate $V$ (whose rows are the vectors), a computer program can automatically verify whether the representations $\{(V,B_y)\}_{y \in \Y}$ are tight.
In particular, if any $(V,B_y)$ is infeasible (i.e. there does not exist any $w$ satisfying all constraints $\langle v_i, w \rangle \leq B_{iy}$), then $V$ is not a $d$-construction.
If $(V,B_y)$ is feasible but some constraints are redundant, i.e. it is not tight, then again we do not have a $d$-construction.

@@

\section{Omitted Proofs}
\begin{proof}[Proof of Theorem~\ref{thm:abs-5outcome-not-2embed}]
By Corollary~\ref{cor:d-embed-implies-d-construction}, we can show there is no $2$-construction for one of the level sets.
We choose the abstain cell, which is $\{p : \max_y p_y \leq 0.5\}$.
Because this equals $\{p : (\forall y) 2p_y \geq \sum_{y'} p_{y'}$\}, we can write the level set as $\{p : Bp \geq 0\}$ for
  \[ B = \begin{pmatrix}
       -1 &  1 &  1 & \cdots & 1  \\
       1  & -1 &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -1
     \end{pmatrix} .
  \]
In other words, $B := \ones - 2I$, where $\ones$ is the all-ones matrix and $I$ is the identity.
In our example, $|\Y| = 5$, so $B$ has width $5$ because there are $5$ outcomes and height $5$ because there are $5$ constraints (one per outcome).

>>>>

\begin{corollary}\label{cor:d-embed-implies-d-construction}
  If a property is $d$-embeddable, then each of its level sets has a $d$-construction.
\end{corollary}

This provides the following recipe for both constructing losses and proving lower bounds.
First, take any level set and represent it as $\{p : Bp \geq 0\}$.
Now if $B$ has $k$ rows, search for a collection of $k$ vectors in $\reals^d$.
For any candidate $V$ (whose rows are the vectors), a computer program can automatically verify whether the representations $\{(V,B_y)\}_{y \in \Y}$ are tight.
In particular, if any $(V,B_y)$ is infeasible (i.e. there does not exist any $w$ satisfying all constraints $\langle v_i, w \rangle \leq B_{iy}$), then $V$ is not a $d$-construction.
If $(V,B_y)$ is feasible but some constraints are redundant, i.e. it is not tight, then again we do not have a $d$-construction.

@@

\section{Higher dimensions}
\begin{proof}[Proof of Corollary~\ref{cor:d-embed-implies-d-construction}]
  Suppose the level set is represented as $\Gamma_r = \{p : Bp \geq 0\}$ with $B \in \reals^{k \times n}$ for the minimum $k$ (i.e. the number of facets of the polytope).
  Let $u \in \reals^d$ be its embedding point.
  The convex expected loss is minimized at $u$ if and only if it has the zero vector as a subgradient at $u$, i.e.
    \[ p \in \Gamma_r \qquad \iff \qquad 0 \in \partial \langle L(u), p \rangle . \]
  Subgradient sets satisfy the following two properties: $\partial(f_1 + f_2) = \partial f_1 + \partial f_2$ where the right side is the Minkowski sum of the subgradient sets (i.e. the set of sums of all pairs); and $\partial(\alpha f_1) = \alpha \partial f_1$.
  So
    \[ \partial \langle p , L(u) \rangle = \sum_{y \in \Y} p_y \partial L(u)_y \]
  where the right side is a Minkowski sum of subgradient sets.
  We now consider the halfspaces defining this set.
  Define the $y$th support function as $H_y(v) = \max_{w \in \partial L(u)_y} \langle w, v \rangle$ to obtain that $w \in \partial \langle p, L(u) \rangle$ if and only if, for all unit vectors $v$,
   \[ \langle v, w \rangle \leq \sum_{y \in \Y} p_y H_y(v) . \]
  In particular, the embedding point is optimal for $p$ if and only if this is satisfied for $w=0$.
  So, letting $B_i$ be the $i$th column of $B$, we have
    \[ \Gamma_r = \{p : \sum_y p_y B_y \geq 0\} = \{p : \sum_y p_y H_y(v) \geq 0 ~ (\forall v, \|v\|=1)\} . \]
  These sets are equal and full-dimensional as polytopes.
  By standard results (e.g.~\cite{grunbaum2013convex}), their H-representations are unique, so we must have $k$ vectors $v_1,\dots,v_k \in \reals^d$ such that $H_y(v_i) = B_{iy}$.
  Collected as rows of $V$, we obtain that each $(V,B_y)$ is a tight halfspace representation, as it contains $\partial L(u)_y$ and $H_y(v_i) = B_{iy}$, meaning there exists $w \in \partial L(u)_y$ such that $\langle w, v_i \rangle = B_{iy}$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:abs-5outcome-not-2embed}]
By Corollary~\ref{cor:d-embed-implies-d-construction}, we can show there is no $2$-construction for one of the level sets.
We choose the abstain cell, which is $\{p : \max_y p_y \leq 0.5\}$.
Because this equals $\{p : (\forall y) 2p_y \geq \sum_{y'} p_{y'}$\}, we can write the level set as $\{p : Bp \geq 0\}$ for
  \[ B = \begin{pmatrix}
       -1 &  1 &  1 & \cdots & 1  \\
       1  & -1 &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -1
     \end{pmatrix} .
  \]
In other words, $B := \ones - 2I$, where $\ones$ is the all-ones matrix and $I$ is the identity.
In our example, $|\Y| = 5$, so $B$ has width $5$ because there are $5$ outcomes and height $5$ because there are $5$ constraints (one per outcome).
